Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-22 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex
add experiments
Fixes

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-examples-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number.
If the examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in V \times
\{1,2,\dots,K\}$ are strongly linearly separable with margin $\gamma$ and 
%%%%%
$\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$
%for all $t \in \cbr{1,\ldots,T}$ satisfy $\norm{x_t} \le R$ 
%%%%%
then the expected number
of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

@@

\begin{theorem}[Mistake lower bound]
\label{theorem:strongly-separable-examples-mistake-lower-bound}
Let $\gamma$ be a positive real number, and $R$ be a non-negative real
number. 
%%%%%
Any (possibly randomized) algorithm makes at least 
$((K-1)/2)\left\lfloor (R/\gamma)^2/4 \right\rfloor$ mistakes in expectation 
on some \textsc{Online
Multiclass Classification with Bandit Feedback} problem with $K \le
(R/\gamma)^2$ classes and labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in V \times
\{1,2,\dots,K\}$ for some inner product space $(V,
\ip{\cdot}{\cdot})$ that are strongly linearly separable with margin $\gamma$ and satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ .
%
%For any (possibly randomized) algorithm $\calA$ for the \textsc{Online
%Multiclass Classification with Bandit Feedback} problem with $K \le
%(R/\gamma)^2$ classes, there exists an inner product space $(V,
%\ip{\cdot}{\cdot})$, a non-negative integer $T$ and a sequence of labeled
%examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ in $V \times
%\{1,2,\dots,K\}$ that are strongly linearly separable with margin $\gamma$, the
%norms satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ and the expected
%number of mistakes $\calA$ makes is at least $\frac{K-1}{2} \left\lfloor
%\frac{1}{4} (R/\gamma)^2 \right\rfloor$.
%%%%%
\end{theorem}

\paragraph{Remark.} 
%%%%%
If $\gamma \le R$ then, irrespective 
%%%%%
of any other conditions on $K$, $R$, and $\gamma$, 
a trivial lower bound on the expected number of mistakes of any randomized
algorithm is $(K-1)/2$. To see this, note that the adversary can choose an
example $(R e_1, y)$, where $e_1$ is some arbitrary unit vector in $V$ and $y$ is a label chosen uniformly from
$\cbr{1,2,\dots,K}$, and show this example $K$ times. The sequence of examples
trivially satisfies the strong linear separability condition, and the
$(K-1)/2$ expected mistake lower bound follows from ~\citep[][Claim 2]{Daniely-Helbertal-2013}. However, it is unclear if the trivial lower
bound is the best possible if $K$ is large, i.e., $\omega((R/\gamma)^2)$. We
leave this as an open problem.

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} can
be extended to nonlinear classification using
\emph{positive definite kernels} (or \emph{kernels}, for short), which are 
functions of the form $k: X^2 \to V$ for some set $X$ and inner product space 
$(V, \ip{\cdot}{\cdot})$ such that the matrix $\left(k(x_i,x_j)\right)_{i,j=1}^m$ 
is positive semidefinite for any positive integer $m$ and $x_1, \dots, x_m \in X$
~\citep[][Definition 2.5]{Scholkopf-Smola-2002}. 
%kernels~\cite{Scholkopf-Smola-2002, Shawe-Taylor-Cristianini-2004}. The formal
%description of the kernelized algorithm is given in
%Algorithm~\ref{algorithm:kernelized}. 
As opposed to explicitly maintaining the
weight vector for each class, the algorithm maintains the set of example-scalar
pairs corresponding to the updates of the non-kernelized algorithm. As a direct
consequence of
Theorem~\ref{theorem:strongly-separable-examples-mistake-upper-bound} we get a
mistake bound for the kernelized algorithm.

\begin{theorem}[Mistake upper bound for kernelized algorithm]
\label{theorem:kernelized-upper-bound}
Let $X$ be a non-empty set, let $(V, \ip{\cdot}{\cdot})$ be an inner product
space. Let $\phi:X \to V$ be a feature map and let $k:X \times X \to \R$,
$k(x,x') = \ip{\phi(x)}{\phi(x')}$ be the associated positive definite
kernel. Let $K$ be a positive integer, $\gamma$ be a positive real number, $R$
be a non-negative real number. If $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in
X \times \{1,2,\dots,K\}$ are labeled examples such that:
\begin{enumerate}
\item the mapped examples $(\phi(x_1), y_1)$, $(\phi(x_2), y_2)$, $\dots$, $(\phi(x_T), y_T)$
are strongly linearly separable with margin $\gamma$,
\item for all $t$ in $\cbr{1,\ldots,T}$, $k(x_t,x_t) \le R^2$,
\end{enumerate}
then the expected number of mistakes Algorithm~\ref{algorithm:kernelized} makes
is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

>>>>

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-examples-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number.
If the examples $(x_1, y_1), \dots, (x_T, y_T) \in V \times \{1,2,\dots,K\}$ are
strongly linearly separable with margin $\gamma$ and $\norm{x_1}, \norm{x_2},
\dots, \norm{x_T} \le R$ then the expected number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

@@

\begin{theorem}[Mistake lower bound]
\label{theorem:strongly-separable-examples-mistake-lower-bound}
Let $\gamma$ be a positive real number, $R$ be a non-negative real number and
let $K \le (R/\gamma)^2$ be a positive integer. Any (possibly randomized)
algorithm makes at least $((K-1)/2)\left\lfloor (R/\gamma)^2/4 \right\rfloor$
mistakes in expectation on some \textsc{Online Multiclass Classification with
Bandit Feedback} problem with $K$ classes and labeled examples $(x_1, y_1),
(x_2, y_2), \dots, (x_T, y_T) \in V \times \{1,2,\dots,K\}$ for some inner
product space $(V, \ip{\cdot}{\cdot})$ that are strongly linearly separable with
margin $\gamma$ and satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$.
\end{theorem}

\paragraph{Remark.}
If $\gamma \le R$ then, irrespective of any other conditions on $K$, $R$, and
$\gamma$, a trivial lower bound on the expected number of mistakes of any
randomized algorithm is $(K-1)/2$. To see this, note that the adversary can
choose an example $(R e_1, y)$, where $e_1$ is some arbitrary unit vector in $V$
and $y$ is a label chosen uniformly from $\cbr{1,2,\dots,K}$, and show this
example $K$ times. The sequence of examples trivially satisfies the strong
linear separability condition, and the $(K-1)/2$ expected mistake lower bound
follows from ~\citep[][Claim 2]{Daniely-Helbertal-2013}. However, it is unclear
if the trivial lower bound is the best possible if $K$ is large, i.e.,
$\omega((R/\gamma)^2)$. We leave this as an open problem.

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} can
be extended to nonlinear classification using \emph{positive definite kernels}
(or \emph{kernels}, for short), which are functions of the form $k: X \times X
\to V$ for some set $X$ and inner product space $(V, \ip{\cdot}{\cdot})$ such
that the matrix $\left(k(x_i,x_j)\right)_{i,j=1}^m$ is a symmetric positive
semidefinite for any positive integer $m$ and $x_1, x_2, \dots, x_m \in X$
\citep[Definition 2.5]{Scholkopf-Smola-2002}. As opposed to explicitly
maintaining the weight vector for each class, the algorithm maintains the set of
example-scalar pairs corresponding to the updates of the non-kernelized
algorithm. As a direct consequence of
Theorem~\ref{theorem:strongly-separable-examples-mistake-upper-bound} we get a
mistake bound for the kernelized algorithm.

\begin{theorem}[Mistake upper bound for kernelized algorithm]
\label{theorem:kernelized-upper-bound}
Let $X$ be a non-empty set, let $(V, \ip{\cdot}{\cdot})$ be an inner product
space. Let $\phi:X \to V$ be a feature map and let $k:X \times X \to \R$,
$k(x,x') = \ip{\phi(x)}{\phi(x')}$ be the associated positive definite
kernel. Let $K$ be a positive integer, $\gamma$ be a positive real number, $R$
be a non-negative real number. If $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in
X \times \{1,2,\dots,K\}$ are labeled examples such that:
\begin{enumerate}
\item the mapped examples $(\phi(x_1), y_1)$, $\dots$, $(\phi(x_T), y_T)$
are strongly linearly separable with margin $\gamma$,
\item for all $t$ in $\cbr{1,\ldots,T}$, $k(x_t,x_t) \le R^2$,
\end{enumerate}
then the expected number of mistakes Algorithm~\ref{algorithm:kernelized} makes
is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

### 2019-01-22 ICML-submission/appendix-experiment.tex
Fixes

\section{Supplementary Materials for Section~\ref{sec:experiments}}
\label{sec:supp-to-experiment}
The three-class data we used in Section~\ref{sec:experiments} is originally generated in a $d=2$ plate as shown in the following figure: 
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}  
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_points}
        \caption{Strongly separable case}
        \label{fig:strong-points}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth} 
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Data distribution for Section~\ref{sec:experiments}}
    % \caption{Comparison between our algorithm and Banditron with different exploration parameter $\epsilon$ under $\gamma=0.05$ and $K=3$. }
\end{figure}

>>>>

\section{Supplementary Materials for Section~\ref{sec:experiments}}
\label{sec:supp-to-experiment}

### 2019-01-22 ICML-submission/appendix-multiclass-perceptron.tex
Fixes

We first show that the set of examples $(x_1,y_1)$, $\ldots$,
$(x_T, y_T)$ we have constructed is weakly linearly separable
with margin $\gamma$. To prove that, we demonstrate vectors $w_1, w_2, \dots, w_K$
satisfying conditions \eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}. We define
$$
w_i = \frac{\gamma}{R} \sum_{\substack{t : 1 \le t \le T \\ y_t = i}} e_t \qquad \text{for $i=1,2,\dots,K$.}
$$
Let $a_i = |\{ t ~:~ 1 \le t \le T, \ y_t = i \}|$ be the number of occurrences of label $i$.
It is easy to see that
$$
\norm{w_i}^2 = \frac{\gamma^2}{R^2} \sum_{\substack{t : 1 \le t \le T \\ y_t = i}} \norm{e_t}^2 = \frac{a_i \gamma^2}{R^2} \qquad \text{for $i=1,2,\dots,K$.}
$$
Since $\sum_{i=1}^K a_i = T$,
$\sum_{i=1}^K \norm{w_i}^2 = T \cdot \frac{\gamma^2}{R^2} \leq 1$, i.e.
the condition
\eqref{equation:weak-linear-separability-1} holds. To verify condition
\eqref{equation:weak-linear-separability-2} consider any labeled example $(x_t,
y_t)$. Then, for any $i$ in $\cbr{1,\ldots,K}$, by the definition of $w_i$, we have
\begin{align*}
\ip{w_i}{x_t}
& = \frac{\gamma}{R} \sum_{\substack{s : 1 \le s \le T \\ y_s = i}} \ip{e_s}{R e_t} \\
& = \gamma \cdot \sum_{\substack{s : 1 \le s \le T \\ y_s = i}} \one[s = t] \\
& = \gamma \cdot \one[y_t = i]
\; .
\end{align*}
Therefore, if $i = y_t$, $\ip{w_i}{x_t} = \gamma$;
otherwise $i \neq y_t$, in which case $\ip{w_i}{x_t} = 0$.
Hence, condition \eqref{equation:weak-linear-separability-2} holds.

We now give a lower bound on the number of mistakes $\calA$ makes.
Define a filtration $\cbr{\calB_t}_{t=0}^T$,
where $\calB_t = \sigma((x_1, y_1, \hat{y}_1), \ldots, (x_t, y_t, \hat{y}_t))$.
As $y_t$ is chosen uniformly from $\cbr{1,\ldots,K}$ and independent of
$\calB_{t-1}$ and $\hat{y}_t$,
\[ \Exp[ \indicator{\widehat y_t \neq y_t} | \calB_{t-1} ] \geq 1 - \frac 1 K \geq \frac 1 2. \]
This implies that
\[ \Exp[ \indicator{\widehat y_t \neq y_t} ] \geq \frac 1 2. \]
Summing over all $t$ in $\cbr{1,\ldots,T}$, we conclude that
\[ \Exp \sbr{ \sum_{t=1}^T \indicator{\widehat y_t \neq y_t} } \geq \frac T 2 = \frac 1 2 \lfloor (R/\gamma)^2 \rfloor, \]
which completes the proof.
\end{proof}

>>>>

We first show that the set of examples $(x_1,y_1)$, $\ldots$,
$(x_T, y_T)$ we have constructed is weakly linearly separable
with margin $\gamma$. To prove that, we demonstrate vectors $w_1, w_2, \dots, w_K$
satisfying conditions \eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}. We define
$$
w_i = \frac{\gamma}{R} \sum_{\substack{t : 1 \le t \le T \\ y_t = i}} e_t \qquad \text{for $i=1,2,\dots,K$.}
$$
Let $a_i = |\{ t ~:~ 1 \le t \le T, \ y_t = i \}|$ be the number of occurrences of label $i$.
It is easy to see that
$$
\norm{w_i}^2 = \frac{\gamma^2}{R^2} \sum_{\substack{t : 1 \le t \le T \\ y_t = i}} \norm{e_t}^2 = \frac{a_i \gamma^2}{R^2} \qquad \text{for $i=1,2,\dots,K$.}
$$
Since $\sum_{i=1}^K a_i = T$,
$\sum_{i=1}^K \norm{w_i}^2 = T \cdot \frac{\gamma^2}{R^2} \leq 1$, i.e.
the condition
\eqref{equation:weak-linear-separability-1} holds. To verify condition
\eqref{equation:weak-linear-separability-2} consider any labeled example $(x_t,
y_t)$. Then, for any $i$ in $\cbr{1,\ldots,K}$, by the definition of $w_i$, we have
\begin{align*}
\ip{w_i}{x_t}
& = \frac{\gamma}{R} \sum_{\substack{s : 1 \le s \le T \\ y_s = i}} \ip{e_s}{R e_t} \\
& = \gamma \cdot \sum_{\substack{s : 1 \le s \le T \\ y_s = i}} \indicator{s = t} \\
& = \gamma \cdot \indicator{y_t = i}
\; .
\end{align*}
Therefore, if $i = y_t$, $\ip{w_i}{x_t} = \gamma$;
otherwise $i \neq y_t$, in which case $\ip{w_i}{x_t} = 0$.
Hence, condition \eqref{equation:weak-linear-separability-2} holds.

We now give a lower bound on the number of mistakes $\calA$ makes.
As $y_t$ is chosen uniformly from $\{1,2,\dots,K\}$ and independent of
$\calA$'s randomization
$$
\Exp[ \indicator{\widehat y_t \neq y_t} ] \ge 1 - \frac{1}{K} \ge \frac{1}{2} \; .
$$
Summing over all $t$ in $\cbr{1,\ldots,T}$, we conclude that
$$
\Exp \sbr{ \sum_{t=1}^T \indicator{\widehat y_t \neq y_t} } \geq \frac T 2 = \frac 1 2 \lfloor (R/\gamma)^2 \rfloor,
$$
which completes the proof.
\end{proof}

### 2019-01-22 ICML-submission/appendix-proofs-for-strongly-separable-examples.tex
Fixes

%$\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$
%$A
%= \sum_{t ~:~ S_t \neq \emptyset} z_t$
%B = \sum_{t ~:~ S_t = \emptyset} z_t
%\sum_{t ~:~ S_t = \emptyset}
\begin{proof}[Proof of \autoref{theorem:strongly-separable-examples-mistake-upper-bound}]
Let $M = \sum_{t=1}^T z_t$ be the number of
mistakes Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} makes.
Let $A = \sum_{t=1}^T \one[S_t \neq \emptyset] z_t$ be the number of mistakes in the rounds
when $S_t \neq \emptyset$, i.e. the number of rounds line~\ref{line:neg-update} is
executed.
In addition, let $B = \sum_{t=1}^T \one[S_t = \emptyset] z_t$ be the
number of mistakes in the rounds when $S_t = \emptyset$.
It can be easily seen that $M = A + B$.

@@

With probability one, the norm of each example is exactly $R$.
We show that with probability
one, the examples are strongly separable with margin $\gamma$. To see
that, consider $w_1^*, w_2^*, \dots, w_K^* \in V$ defined by
$$
w_i^* = \sqrt{2} \frac{\gamma}{R} \left( \sum_{j ~:~ \ell_j = i} e_j \right) - \frac{\sqrt{2}}2 \frac{\gamma}{R} e_{M+1}
$$
for $i=1,2,\dots,K$.

for $i \in \{1,2,\dots,K\}$ and $j \in
\{1,2,\dots,M\}$, we consider the inner product of $w_i^*$ and $v_j$.
If $i = \ell_j$, $\ip{w_i^*}{v_j} = \gamma - \frac \gamma 2 = \frac \gamma 2$;
otherwise $i \neq \ell_j$, in which case
$\ip{w_i^*}{v_j} = 0 - \frac \gamma 2 = - \frac \gamma 2$.
This means that $w_1^*, w_2^*, \dots, w_K^*$ satisfy
conditions
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}. Condition \eqref{equation:strong-linear-separability-1}
is satisfied since
\begin{multline*}
\sum_{i=1}^K \norm{w_i^*}^2
= 2 \frac{\gamma^2}{R^2} \sum_{j=1}^M \norm{e_j}^2 +  \frac{\gamma^2}{2 R^2} K \norm{e_{M+1}}^2 \\
= 2 \frac{\gamma^2}{R^2} M + \frac{\gamma^2}{2 R^2} K
\le \frac{1}{2} + \frac{1}{2}
= 1 \; .
\end{multline*}

@@

Define a filtration $\cbr{\calB_j}_{j=0}^M$,
where $\calB_j = \sigma((x_1, y_1, \hat{y}_1), \ldots, (x_{(K-1)j}, y_{(K-1)j}, \hat{y}_{(K-1)j}))$
for every $j$ in $\{1,2,\dots,M\}$.
By Claim 2 of~\citet{Daniely-Helbertal-2013}, as $\ell_j$
is chosen uniformly from $\cbr{1,\dots,K}$ and independent of $\calB_{j-1}$ and $\calA$'s
randomness,
$$
\Exp \sbr{ \sum_{t=(K-1)(j-1) + 1}^{(K-1)j} z_t ~\Bigg|~ \calB_{j-1} } \geq \frac{K-1}2 \; .
$$
This implies that
$$
\Exp \sbr{ \sum_{t=(K-1)(j-1) + 1}^{(K-1)j} z_t } \ge \frac{K-1}2 \; .
$$
Summing over all $j$ in $\{1,2,\dots,M\}$,
$$
\Exp \sbr{ \sum_{t=1}^{(K-1)M} z_t} \geq \frac{K-1}2 \cdot M = \frac{K-1}2 \left\lfloor \frac 1 4 (R/\gamma)^2 \right\rfloor \; .
$$

%In these rounds,
%the algorithm is guessing the label $\ell_j$. Since $\ell_j$ is chosen uniformly
%at random from the set $\{1,2,\dots,K\}$ and feedback is only binary, the
%expected number of mistakes the algorithm makes in these rounds is at least
%$\frac{K-1}{2}$. Altogether, the algorithm makes at least $\frac{K-1}{2} M$
%mistakes in expectation.

%\chicheng{I think this part need to be formalized, by directly using
%~\cite{Daniely-Helbertal-2013}, Claim 2.}
%Can we assume without loss of generality
%that a deterministic prediction algorithm will have a smaller expected cost?
%If so, define $F(k)$ be the number of mistake that will be made by the algorithm,
%if the adversary commits to a label uniformly at random from $\cbr{1,2,\ldots,k}$,
%and the guessing of the label lasts for $k$ rounds.
%we can construct a recurrence of $F$: $F(k) = F(k-1) + \frac{k-1}{k}$, and $F(1) = 0$.
%Therefore, $F(K) = \Omega(K)$.

>>>>

\begin{proof}[Proof of \autoref{theorem:strongly-separable-examples-mistake-upper-bound}]
Let $M = \sum_{t=1}^T z_t$ be the number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes. Let $A = \sum_{t=1}^T \one[S_t \neq \emptyset] z_t$ be the number of
mistakes in the rounds when $S_t \neq \emptyset$, i.e. the number of rounds
line~\ref{line:neg-update} is executed. In addition, let $B = \sum_{t=1}^T
\one[S_t = \emptyset] z_t$ be the number of mistakes in the rounds when $S_t =
\emptyset$. It can be easily seen that $M = A + B$.

@@

The norm of each example is exactly $R$. The examples are strongly separable
with margin $\gamma$. To see that, consider $w_1^*, w_2^*, \dots, w_K^* \in V$
defined by
$$
w_i^* = \sqrt{2} \frac{\gamma}{R} \left( \sum_{j ~:~ \ell_j = i} e_j \right) - \frac{\sqrt{2}}2 \frac{\gamma}{R} e_{M+1}
$$
for $i=1,2,\dots,K$.

For $i \in \{1,2,\dots,K\}$ and $j \in
\{1,2,\dots,M\}$, consider the inner product of $w_i^*$ and $v_j$.
If $i = \ell_j$, $\ip{w_i^*}{v_j} = \gamma - \frac \gamma 2 = \frac \gamma 2$;
otherwise $i \neq \ell_j$, in which case
$\ip{w_i^*}{v_j} = 0 - \frac \gamma 2 = - \frac \gamma 2$.
This means that $w_1^*, w_2^*, \dots, w_K^*$ satisfy
conditions
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}. Condition \eqref{equation:strong-linear-separability-1}
is satisfied since
\begin{multline*}
\sum_{i=1}^K \norm{w_i^*}^2
= 2 \frac{\gamma^2}{R^2} \sum_{j=1}^M \norm{e_j}^2 +  \frac{\gamma^2}{2 R^2} K \norm{e_{M+1}}^2 \\
= 2 \frac{\gamma^2}{R^2} M + \frac{\gamma^2}{2 R^2} K
\le \frac{1}{2} + \frac{1}{2}
= 1 \; .
\end{multline*}

@@

Define a filtration of $\sigma$-algebras $\cbr{\calB_j}_{j=0}^M$, where $\calB_j
= \sigma((x_1, y_1, \hat{y}_1), \ldots, (x_{(K-1)j}, y_{(K-1)j},
\hat{y}_{(K-1)j}))$ for every $j$ in $\{1,2,\dots,M\}$. By Claim 2
of~\citet{Daniely-Helbertal-2013}, as $\ell_j$ is chosen uniformly from
$\cbr{1,\dots,K}$ and independent of $\calB_{j-1}$ and $\calA$'s randomness,
$$
\Exp \sbr{ \sum_{t=(K-1)(j-1) + 1}^{(K-1)j} z_t ~\Bigg|~ \calB_{j-1} } \ge \frac{K-1}{2} \; .
$$
This implies that
$$
\Exp \sbr{ \sum_{t=(K-1)(j-1) + 1}^{(K-1)j} z_t } \ge \frac{K-1}{2} \; .
$$
Summing over all $j$ in $\{1,2,\dots,M\}$,
$$
\Exp \sbr{ \sum_{t=1}^{(K-1)M} z_t} \geq \frac{K-1}2 \cdot M = \frac{K-1}2 \left\lfloor \frac 1 4 (R/\gamma)^2 \right\rfloor \; .
$$

### 2019-01-22 ICML-submission/empirical-comparison.tex
Fixes

 \section{Experiments}
\label{sec:experiments}
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.23\textwidth} 
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0.1cm, 0, 0}, clip]{figures/strong3}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth} 
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0.1cm, 0, 0}, clip]{figures/weak3}
        \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Comparison between our algorithm and Banditron with different exploration parameter $\epsilon$ under $\gamma=0.05$ and $K=3$. }
    %\includegraphics[width=0.25\textwidth]{figures/weak3}
    \label{fig:both}
\end{figure}

We compare our algorithm with \textsc{Banditron} empirically to show that our kernelized bandit algorithm can have significant advantage over \textsc{Banditron} is the separable case.

We generate strongly and weakly separable samples with $K=3$ in a similar sense those in Figure~\ref{figure:weakly-linearly-separable-examples-with-margin}. They are originally generated in a $d=2$ plate, and then lifted to $d=3$ by adding a constant feature. We provide the figures of their exact look in Appendix~\ref{sec:supp-to-experiment}. Different from Figure~\ref{figure:weakly-linearly-separable-examples-with-margin}, the three classes are non-uniform and imbanlanced: class 1 occupies a smaller sector between $-15^\circ$ and $15^\circ$, while class 2 and class 3 occupy $[15^\circ, 180^\circ]$ and $[-180^\circ, -15^\circ]$ respectively. On the other hand, $80\%$ of data comes from class 1, while $10\%$ comes from class 2 and 3 each.

In real world, data can often be non-uniform and imbalanced. We demonstrate that \textsc{Banditron} is weak especially under these situations, while our algorithm still performs well.

We choose sample size as $5\times 10^6$ and margin as $\gamma=0.05$ (in both strongly weakly separable cases). All results are averaged over $20$ runs.

We plot the number of mistakes against time in Figure~\ref{fig:both}. We can see there is indeed a dilemma for \textsc{Banditron}: when picking a relatively large exploration rate, it suffers from the cost of continuous exploration; when picking a relatively small exploration rate, it cannot adapt its model quickly enough (this problem exacerbates when a majority class occupies a small region, like class 1). Striking a balance between the two extremes leads to the $\sqrt{T}$ mistake bound. On the contrary, our algorithm makes significant progress on every mistake it makes, leading to a finite mistake bound.

>>>>

\section{Experiments}
\label{sec:experiments}

We compare our algorithm with \textsc{Banditron} empirically to show that our
kernelized bandit algorithm can have significant advantage over
\textsc{Banditron} in the separable case.

We generated strongly and weakly separable samples with $K=3$ classes shown
in~\autoref{figure:strongly-and-weakly-separable-datasets}. The data sets are
shown as two-dimensional, however, we used a third dimension (so-called
\emph{constant} feature) i.e. a feature that for every example is set to $1$. We
added the constant feature so that \autoref{figure:strongly-separable-dataset}
is actually strongly lineraly separable. For both data sets, the sample size is
$T=5\times 10^6$ and the margin is $\gamma=0.05$.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\begin{center}
\hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_points}
\caption{Strongly separable case}
\label{figure:strongly-separable-dataset}
\end{center}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\centering
\hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_points}
\caption{Weakly separable case}
\label{figure:weakly-separable-dataset}
\end{subfigure}
\vspace*{-0.2cm}
\caption{Strongly and weakly generated data sets with $K=3$ classes. All examples lie in
the unit ball. Class 1 is depicted red. Classes 2 and 3 are depcited gree and
blue, respectively. $80\%$ of example belong to class 1, $10\%$ of the data
belong to class 2 and $10\%$ belong to class 3. Class 1 occupies the angle
interval $[-15^\circ, 15^\circ]$, while classes 2 and 3 occupy angle intervals
$[15^\circ, 180^\circ]$ and $[-180^\circ, -15^\circ]$ respectively. We removed
points that lie within margin $\gamma=0.05$ of the linear separators.}
\label{figure:strongly-and-weakly-separable-datasets}
\end{figure}

All results are averaged over $20$ runs.
We plot the number of mistakes against time in Figure~\ref{fig:both}. We can see
there is indeed a dilemma for \textsc{Banditron}: when picking a relatively
large exploration rate, it suffers from the cost of continuous exploration; when
picking a relatively small exploration rate, it cannot adapt its model quickly
enough (this problem exacerbates when a majority class occupies a small region,
like class 1). Striking a balance between the two extremes leads to the
$\sqrt{T}$ mistake bound. On the contrary, our algorithm makes significant
progress on every mistake it makes, leading to a finite mistake bound.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0.1cm, 0, 0}, clip]{figures/strong3}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0.1cm, 0, 0}, clip]{figures/weak3}
        \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Comparison between our algorithm and Banditron with different exploration parameter $\epsilon$ under $\gamma=0.05$ and $K=3$.}
    \label{fig:both}
\end{figure}
